Adversarial Variational Optimization of\\ Non-Differentiable Simulators

### 2018-02-06 icml/avo.tex
save

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{XXXXX}

@@

\begin{abstract}
    Complex computer simulators are increasingly used across fields
of science as generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often difficult, as
simulators rarely admit a tractable density or likelihood function. We introduce
Adversarial Variational Optimization (AVO), a likelihood-free inference
algorithm for fitting a non-differentiable generative model incorporating ideas
from generative adversarial networks (GANs) and variational optimization. We adapt the
training procedure of Wasserstein GANs by replacing the
differentiable generative network with a domain-specific simulator. We solve the
resulting non-differentiable minimax problem by minimizing variational upper
bounds of the two adversarial objectives.
Effectively, the procedure results in
learning a proposal distribution over simulator parameters, such that
the Wasserstein distance between the marginal distribution of the synthetic data and
the empirical distribution of observed data is minimized.
We present results of the method with simulators producing both
discrete and continuous data.
\end{abstract}

@@

In this work, we develop a likelihood-free inference algorithm
for non-differentiable, implicit generative models.
We adapt the adversarial
training procedure of generative adversarial
networks~\cite{goodfellow2014generative} by replacing the implicit generative
network with a domain-based scientific simulator, and solve the resulting
non-differentiable minimax problem by minimizing variational upper
bounds~\cite{2011arXiv1106.4487W,2012arXiv1212.4507S} of the adversarial
objectives.
The objective of the algorithm is to match the marginal distribution of
the generated data to the empirical distribution of the observations.

@@

This work has also many connections to work on variational inference, in which
the goal is to optimize the recognition model $q(\bfz, \bftheta|\bfpsi)$ so that
it is close to the true posterior $p(\bfz, \bftheta |\mathbf{x})$. There have
been efforts to extend variational inference to intractable likelihoods;
however, many require restrictive assumptions.  In \cite{tran2017variational},
the authors consider Variational Bayes with an Intractable Likelihood (VBIL). In
that approach ``the only requirement is that the intractable likelihood can be
estimated unbiasedly.'' In the case of simulators, they propose to use the
ABC-likelihood with an $\epsilon$-kernel. The ABC likelihood is only unbiased as
$\epsilon \to 0$,   thus this method inherits the drawbacks of the
ABC-likelihood including the choice of summary statistics and the inefficiency
in evaluating the ABC likelihood for high-dimensional data and small $\epsilon$.
More recently, \cite{2017arXiv170208896T} adapted variational inference to
hierarchical implicit models defined on simulators. In this work, the authors
step around the intractable likelihoods by reformulating the optimization of the
ELBO in terms of a neural and differentiable approximation $r$ of the log-likelihood
ratio $\log \tfrac{p}{q}$, thereby effectively using the same core principle as used
in GANs~\cite{2016arXiv161003483M}.

>>>>

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{likelihood-free inference, implicit models, generative adversarial networks}

@@

\begin{abstract}
Complex computer simulators are increasingly used across fields
of science as generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often difficult, as
simulators rarely admit a tractable density or likelihood function. We introduce
Adversarial Variational Optimization (AVO), a likelihood-free inference
algorithm for fitting a non-differentiable generative model incorporating ideas
from generative adversarial networks, variational optimization and empirical Bayes. We adapt the
training procedure of Wasserstein GANs by replacing the
differentiable generative network with a domain-specific simulator. We solve the
resulting non-differentiable minimax problem by minimizing variational upper
bounds of the two adversarial objectives.
Effectively, the procedure results in
learning a proposal distribution over simulator parameters, such that
the Wasserstein distance between the marginal distribution of the synthetic data and
the empirical distribution of observed data is minimized.
We present results of the method with simulators producing both
discrete and continuous data.
\end{abstract}

@@

In this work, we develop a likelihood-free inference algorithm
for non-differentiable, implicit generative models.
We adapt the adversarial
training procedure of generative adversarial
networks~\cite{goodfellow2014generative} by replacing the implicit generative
network with a domain-based scientific simulator, and solve the resulting
non-differentiable minimax problem by minimizing variational upper
bounds~\cite{2011arXiv1106.4487W,2012arXiv1212.4507S} of the adversarial
objectives.
The objective of the algorithm is to match the marginal distribution of
the synthetic data to the empirical distribution of observations.

@@

This work has also many connections to work on variational inference, in which
the goal is to optimize the recognition model $q(\bfz, \bftheta|\bfpsi)$ so that
it is close to the true posterior $p(\bfz, \bftheta |\mathbf{x})$. There have
been efforts to extend variational inference to intractable likelihoods;
however, many require restrictive assumptions.  In \cite{tran2017variational},
the authors consider Variational Bayes with an Intractable Likelihood (VBIL). In
that approach ``the only requirement is that the intractable likelihood can be
estimated unbiasedly.'' In the case of simulators, they propose to use the
ABC-likelihood with an $\epsilon$-kernel. The ABC likelihood is only unbiased as
$\epsilon \to 0$,   thus this method inherits the drawbacks of the
ABC-likelihood including the choice of summary statistics and the inefficiency
in evaluating the ABC likelihood for high-dimensional data and small $\epsilon$.
More recently, \cite{2017arXiv170208896T} adapted variational inference to
hierarchical implicit models defined on simulators. In this work, the authors
step around the intractable likelihoods by reformulating the optimization of the
ELBO in terms of a neural and differentiable approximation $r$ of the log-likelihood
ratio $\log \tfrac{p}{q}$, thereby effectively using the same core principle as used
in GANs~\cite{2016arXiv161003483M}. With a similar objective, \cite{2017arXiv171203353M}
adapt variational inference to a non-differentiable cardiac simulator by maximizing
the ELBO using Bayesian optimization, hence bypassing altogether the need for gradient estimates.
